---
title: "R Notebook"
output: html_notebook
---

#Non-overlapping gene groups:

Type 1: Genes that are involved only in central or optimal hubs.
           -Type 1 optimal.
           -Type 1 central.
Type 2: Genes that participate  both in central and optimal hubs, but not in hyper hubs.
Type 3: Genes that associate only with hyper hubs.
Type 4: Genes that participate both in hyper hubs and in one of the other hubs (central or optimal),
           -Type 4 hyper-optimal.
           -Type 4 hyper-central.
Type 5: Genes that are in optimal, central and hyper hubs. 


#NEURON ranked by pure genes
```{r fig.height=3, fig.width=14, message=F, warning=FALSE}
library(clusterProfiler)
library(ReactomePA)
setwd("~/AA_PostDoc/human-metnet/results/stimulated_NAMU_results_analysis/functional_analysis_of_hubs")
source("analysis_functions.R")

get_genes_from_each_hub('Neuron_IDs_centrality_optimality.csv', cut_off_optimality = 0.01 , cut_off_centrality =  0.01) -> list_of_genes_by_hubs
get_pure_gene_groups(list_of_genes_by_hubs) -> gene_groups
#para guardar resultados gene_groups
gene_groups %>% tibble( group = names(gene_groups), genes=.) %>% unnest -> neuron_genes_for_xlsx
#analisis
gene_groups %>%compareCluster(fun='enrichPathway', pvalueCutoff=1e-3)  %>% dotplot
gene_groups %>%compareCluster(fun='enrichKEGG', pvalueCutoff=1e-3) %>% dotplot


```
#ASTROCYTE ranked by pure genes
```{r fig.height=3, fig.width=14, message=F, warning=FALSE}



astro_url <- 'https://github.com/DeepenData/human-metnet/blob/main/results/stimulated_NAMU_results_analysis/functional_analysis_of_hubs/Astrocyte_IDs_centrality_optimality.xlsx'



get_genes_from_each_hub(astro_url, cut_off_optimality = 0.01 , cut_off_centrality =  0.01) -> list_of_genes_by_hubs
get_pure_gene_groups(list_of_genes_by_hubs) -> gene_groups
#para Guardar resultados
gene_groups %>% tibble( group = names(gene_groups), genes=.) %>% unnest -> astrocyte_genes_for_xlsx
#analisis
gene_groups %>%compareCluster(fun='enrichPathway', pvalueCutoff=1e-3)  %>% dotplot
gene_groups %>%compareCluster(fun='enrichKEGG', pvalueCutoff=1e-3) %>% dotplot
```

Exportar resultados en .xlsx

```{r}
neuron_genes_for_xlsx %>% writexl::write_xlsx(paste0('AA_neuron_genes','.xlsx'))
astrocyte_genes_for_xlsx %>%writexl::write_xlsx(paste0('AA_astrocyte_genes','.xlsx'))
```



######################--------------------------------------------END------------------------------------------#########################################3
###########################################################



```{r fig.height=3, fig.width=14, message=F, warning=FALSE}
library(DOSE)
library(readxl)
library(clusterProfiler)
library(ReactomePA)
library(tidyverse)

DLL_20210504_merged_lists <- read_excel("DLL_20210504_merged_lists.xlsx", 
    sheet = "All types")

purrr::compose(as.character, na.omit)   -> getrid_NAs_clean
DLL_20210504_merged_lists %>% map(getrid_NAs_clean) -> dll

dll %>%compareCluster(fun='enrichPathway',  organism ="mouse", pvalueCutoff=1e-3) %>% dotplot
dll %>%compareCluster(fun='enrichKEGG',  organism ="mmu", pvalueCutoff=1e-3) %>% dotplot

#enrichDGN
```
#Astrocyte-neuron common genes

```{r fig.height=3, fig.width=14, message=F, warning=FALSE}


get_genes_from_each_hub('Astrocyte_IDs_centrality_optimality.xlsx') -> Astrocyte_genes_by_hubs
get_genes_from_each_hub('Neuron_IDs_centrality_optimality.xlsx') -> Neuron_genes_by_hubs

purrr::compose(unlist,flatten) -> condensed
intersect(condensed(Astrocyte_genes_by_hubs),condensed(Neuron_genes_by_hubs)) -> commom_neu_astro
setdiff(condensed(Astrocyte_genes_by_hubs),condensed(Neuron_genes_by_hubs)) -> exclusive_astrocyte
setdiff(condensed(Neuron_genes_by_hubs),condensed(Astrocyte_genes_by_hubs)) -> exclusive_neuron

map2(Neuron_genes_by_hubs, Astrocyte_genes_by_hubs, union)-> the_union

lapply(the_union, function(a_list_entry) {setdiff(a_list_entry , exclusive_astrocyte) }) -> the_union_minus_astro
lapply(the_union_minus_astro, function(a_list_entry) {setdiff(a_list_entry , exclusive_neuron) }) -> only_commom

condensed(only_commom) %in% commom_neu_astro %>% all
commom_neu_astro   %in% condensed(only_commom)  %>% all

only_commom %>% get_pure_gene_groups -> gene_groups
gene_groups %>% tibble( group = names(gene_groups), genes=.) %>% unnest -> neuron_astrocyte_genes
gene_groups %>%compareCluster(fun='enrichPathway', pvalueCutoff=1e-3)  %>% dotplot
gene_groups %>%compareCluster(fun='enrichKEGG', pvalueCutoff=1e-3) %>% dotplot
```
```{r}
neuron_genes$genes %>% duplicated() %>% all
astrocyte_genes$genes %>% duplicated() %>% all
neuron_astrocyte_genes$genes %>% duplicated() %>% all
```

```{r}
neuron_genes %>% writexl::write_xlsx(paste0('AA_neuron_genes','.xlsx'))
astrocyte_genes %>%writexl::write_xlsx(paste0('AA_astrocyte_genes','.xlsx'))
neuron_astrocyte_genes %>% writexl::write_xlsx(paste0('AA_neuron_astrocyte_genes','.xlsx'))
```






















