---
title: "R Notebook"
output: html_notebook
---


```{r fig.height=3, fig.width=14, message=FALSE, warning=FALSE}
library(clusterProfiler)
library(DOSE)
library(ReactomePA)
library(tidyverse)
library(readxl)
setwd("~/AA_PostDoc/human-metnet/results/stimulated_NAMU_results_analysis/functional_analysis_of_hubs")

c('PGM', 'ACYP', 'PGI', 'PGK','PYK', 'HEX1', 'DPGase', 'TPI', 'PFK', 'ENO', 'GAPD', 'DPGM', 'FBA', 'G3PD2m' ,
'ACYP_Neuron', 'DPGM_Neuron', 'DPGase_Neuron', 'ENO_Neuron', 'FBA_Neuron', 'G3PD2m_Neuron', 'GAPD_Neuron', 'HEX1_Neuron', 'PFK_Neuron', 'PGI_Neuron', 'PGK_Neuron', 'PGM_Neuron', 'PYK_Neuron', 'TPI_Neuron' ,
'ATPS4m_Neuron', 'CYOOm2_Neuron', 'CYOR-u10m_Neuron', 'NADH2-u10m_Neuron', 'PPA_Neuron', 'PPAm_Neuron' ,
'PPAm', 'ATPS4m', 'CYOOm2', 'CYOR-u10m', 'NADH2-u10m', 'PPA', 'GLCt1r', 'GLCt1r_Neuron', 'GLCt2r') -> subsystems
filter_out_subsystems <- function(df){df[!(df$ID %in% subsystems),]}
get_genes_from_module <- function(hubs){     hubs %>% 
                                             dplyr::select(matches('gene')) %>% purrr::reduce(c) %>% na.exclude() %>% str_c(collapse = ", ") %>% 
                                            str_split(pattern=",") %>%str_extract_all('\\d+') %>% unlist() %>% str_trim %>% unique()}
```

### NEURON
```{r fig.height=3, fig.width=14, message=FALSE, warning=FALSE}
cut_off_optimality <-0.01  # 0.01
cut_off_centrality <- 0.01 # 0.01
central_hubs <- readxl::read_excel('Neuron_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Optimality")%>%filter(Centrality > cut_off_centrality)%>% filter_out_subsystems
optimal_hubs <- readxl::read_excel('Neuron_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Centrality")%>%filter(Optimality > cut_off_optimality)%>% filter_out_subsystems
central_hubs %<>%  get_genes_from_module 
optimal_hubs %<>%  get_genes_from_module 
intersect(central_hubs, optimal_hubs) -> hyper_hubs
setdiff(central_hubs,hyper_hubs) -> central_hubs
setdiff(optimal_hubs,hyper_hubs) -> optimal_hubs
list(optimal_hubs, central_hubs, hyper_hubs)  %>% purrr::set_names(c("Optimal", "Central", "Hyper"))  -> hubs
intersect(central_hubs, optimal_hubs)
intersect(central_hubs, hyper_hubs)
intersect(optimal_hubs, hyper_hubs)   
hubs %>% glimpse()
compareCluster(hubs, fun='enrichPathway', pvalueCutoff=1e-3) -> clusters_plot; clusters_plot %>% dotplot
compareCluster(hubs, fun='enrichKEGG', pvalueCutoff=1e-3) %>% dotplot

```
## ASTROCYTE
```{r fig.height=3, fig.width=14, message=FALSE, warning=FALSE}
cut_off_optimality <-0.01  # 0.01
cut_off_centrality <- 0.01 # 0.01
central_hubs <- readxl::read_excel('Astrocyte_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Optimality")%>%filter(Centrality > cut_off_centrality)%>% filter_out_subsystems
optimal_hubs <- readxl::read_excel('Astrocyte_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Centrality")%>%filter(Optimality > cut_off_optimality)%>% filter_out_subsystems


central_hubs %<>%  get_genes_from_module 
optimal_hubs %<>%  get_genes_from_module 
intersect(central_hubs, optimal_hubs) -> hyper_hubs
setdiff(central_hubs,hyper_hubs) -> central_hubs
setdiff(optimal_hubs,hyper_hubs) -> optimal_hubs
list(optimal_hubs, central_hubs, hyper_hubs)  %>% purrr::set_names(c("Optimal", "Central", "Hyper"))  -> hubs
intersect(central_hubs, optimal_hubs)
intersect(central_hubs, hyper_hubs)
intersect(optimal_hubs, hyper_hubs)   
hubs %>% glimpse()
compareCluster(hubs, fun='enrichPathway', pvalueCutoff=1e-3) -> clusters_plot; clusters_plot %>% dotplot
compareCluster(hubs, fun='enrichKEGG', pvalueCutoff=1e-3) %>% dotplot

```


#NEURON BY rxns
```{r fig.height=3, fig.width=14, message=FALSE, warning=FALSE}
cut_off_optimality <-0.01  # 0.01
cut_off_centrality <- 0.01 # 0.01
central_hubs <- readxl::read_excel('Neuron_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Optimality")%>%filter(Centrality > cut_off_centrality)%>% filter_out_subsystems
optimal_hubs <- readxl::read_excel('Neuron_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Centrality")%>%filter(Optimality > cut_off_optimality)%>% filter_out_subsystems

central_optimal_hubs <-  readxl::read_excel('Neuron_IDs_centrality_optimality.xlsx')%>% filter_out_subsystems

intersect(central_hubs$ID, optimal_hubs$ID) -> hyper_IDs
setdiff(central_hubs$ID,hyper_IDs)          -> central_IDs
 setdiff(optimal_hubs$ID,hyper_IDs)         -> optimal_IDs
 
central_hubs         %>% filter(ID %in% central_IDs) %>%  get_genes_from_module -> central_genes
optimal_hubs         %>% filter(ID %in% optimal_IDs) %>%  get_genes_from_module -> optimal_genes
central_optimal_hubs %>% filter(ID %in% hyper_IDs) %>% get_genes_from_module -> hyper_genes



intersect(central_genes,optimal_genes) -> super

intersect(super,hyper_genes)   ->  super_and_hyper

intersect(optimal_genes,hyper_genes)  %>% setdiff(super_and_hyper)-> optimal_and_hyper
intersect(central_genes,hyper_genes)  %>% setdiff(super_and_hyper)-> central_and_hyper

vector_to_substract <-  c(hyper_genes, super,super_and_hyper, optimal_and_hyper,central_and_hyper ) %>% unique()

setdiff(central_genes,vector_to_substract) -> pure_central
setdiff(optimal_genes,vector_to_substract) -> pure_optimal


hubs_rxn_based <- list(pure_optimal, pure_central, super, hyper_genes, super_and_hyper, optimal_and_hyper, central_and_hyper )  %>% 
               purrr::set_names(c("pure_optimal", "pure_central", "super", "hyper", "super_and_hyper",  'optimal_and_hyper', 'central_and_hyper'))  
hubs_rxn_based %>% glimpse()

compareCluster(hubs_rxn_based, fun='enrichPathway', pvalueCutoff=1e-3)  %>% dotplot
compareCluster(hubs_rxn_based, fun='enrichKEGG', pvalueCutoff=1e-3) %>% dotplot



```

ASTROCYTE BY rxns
```{r fig.height=3, fig.width=14, message=FALSE, warning=FALSE}
cut_off_optimality <-0.01  # 0.01
cut_off_centrality <- 0.01 # 0.01
central_hubs <- readxl::read_excel('Astrocyte_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Optimality")%>%filter(Centrality > cut_off_centrality)%>% filter_out_subsystems
optimal_hubs <- readxl::read_excel('Astrocyte_IDs_centrality_optimality.xlsx')%>% dplyr::select(-"Centrality")%>%filter(Optimality > cut_off_optimality)%>% filter_out_subsystems

central_optimal_hubs <-  readxl::read_excel('Astrocyte_IDs_centrality_optimality.xlsx')%>% filter_out_subsystems

intersect(central_hubs$ID, optimal_hubs$ID) -> hyper_IDs
setdiff(central_hubs$ID,hyper_IDs)          -> central_IDs
 setdiff(optimal_hubs$ID,hyper_IDs)         -> optimal_IDs
 
central_hubs         %>% filter(ID %in% central_IDs) %>%  get_genes_from_module -> central_genes
optimal_hubs         %>% filter(ID %in% optimal_IDs) %>%  get_genes_from_module -> optimal_genes
central_optimal_hubs %>% filter(ID %in% hyper_IDs) %>% get_genes_from_module -> hyper_genes



intersect(central_genes,optimal_genes) -> super

intersect(super,hyper_genes)   ->  super_and_hyper

intersect(optimal_genes,hyper_genes)  %>% setdiff(super_and_hyper)-> optimal_and_hyper
intersect(central_genes,hyper_genes)  %>% setdiff(super_and_hyper)-> central_and_hyper

vector_to_substract <-  c(hyper_genes, super,super_and_hyper, optimal_and_hyper,central_and_hyper ) %>% unique()

setdiff(central_genes,vector_to_substract) -> pure_central
setdiff(optimal_genes,vector_to_substract) -> pure_optimal


hubs_rxn_based <- list(pure_optimal, pure_central, super, hyper_genes, super_and_hyper, optimal_and_hyper, central_and_hyper )  %>% 
               purrr::set_names(c("pure_optimal", "pure_central", "super", "hyper", "super_and_hyper",  'optimal_and_hyper', 'central_and_hyper'))  
hubs_rxn_based %>% glimpse()

compareCluster(hubs_rxn_based, fun='enrichPathway', pvalueCutoff=1e-3)  %>% dotplot
compareCluster(hubs_rxn_based, fun='enrichKEGG', pvalueCutoff=1e-3) %>% dotplot



```



```{r}
path_figure_2 = "~/AA_PostDoc/human-metnet/results/stimulated_NAMU_results_analysis/Figure_2/"
readRDS(paste0(path_figure_2, "all_data.rds")) %>% .[["delta_geometric"]] %>% dplyr::select(c(ID,Name, Reaction,`Node type`)) -> all_data
readxl::read_excel('IDs_centrality_optimality.xlsx')  %>% inner_join(all_data) -> IDs_centrality_optimality.xlsx

split_genes <- function(a_string){  if (is.na(a_string)){return("")}
                                  a_string %>% str_split(pattern=",")%>%unlist() %>% str_trim}


split_into_strings <- function(df,column){df[[column]] <- map(df[[column]] , split_genes) %>% paste0 
                                           return(df)}

IDs_centrality_optimality.xlsx %<>% split_into_strings('recon3_genes' ) %>% split_into_strings('Lewis2010_genes' ) %>% split_into_strings('VMH_gene_IDs' )  %>% split_into_strings('KEGG_genes' ) 
get_rxns_from_genes <- function(a_gene){   a_gene1 <- paste0('\"', a_gene ,'\"',collapse = '')
                                           IDs_centrality_optimality.xlsx %>% filter(str_detect(recon3_genes,a_gene1) | str_detect(Lewis2010_genes,a_gene1)| str_detect(VMH_gene_IDs,a_gene1)| str_detect(KEGG_genes,a_gene1)) }#%>%
                                          # dplyr::select(c('ID','Name','Reaction',"Node type"))  %>% mutate(Gene = a_gene)}

map_dfr(hubs$Hyper,get_rxns_from_genes) %>% unique() -> H
map_dfr(hubs$Central,get_rxns_from_genes) %>% unique()->C
map_dfr(hubs$Optimal,get_rxns_from_genes) %>% unique()->O
rbind(H,C,O) -> all_hubs_info
```

ketone bodies:
BDHm    (R)-3-Hydroxybutanoate:NAD+ Oxidoreductase bhb[m] + nad[m] <=> acac[m] + h[m] + nadh[m]
  genes: 56898(BDH2) 622(BDH1)

OCOAT1m 3-Oxoacid Coa-Transferase acac[m] + succoa[m] -> aacoa[m] + succ[m]
 genes: 5019(OXCT1) 64064(OXCT2)


AACOAT  Acetoacetyl Coenzyme A:Acetate Coa-Transferase acac[c] + atp[c] + coa[c] -> aacoa[c] + amp[c] + ppi[c]:
 genes: 65985(AACS)


HMGLm   Hydroxymethylglutaryl Coenzyme A Lyase hmgcoa[m] -> acac[m] + accoa[m]
  genes: 	3155(HMGCL) 54511(HMGCLL1)

ACACT1rm Acetyl Coenzyme A C-Acetyltransferase, Mitochondrial 2.0 accoa[m] -> aacoa[m] + coa[m]

HMGCOASim Hydroxymethylglutaryl Coenzyme A Synthase (Ir) aacoa[m] + accoa[m] + h2o[m] -> coa[m] + h[m] + hmgcoa[m]
  genes: 3157(HMGCS1) 3158(HMGCS2)

CSm Citrate Synthase accoa[m] + h2o[m] + oaa[m] -> cit[m] + coa[m] + h[m]
 genes: 	1431(CS)

ACSm Acetyl Coenzyme A Synthetase ac[m] + atp[m] + coa[m] -> accoa[m] + amp[m] + ppi[m]
	genes: 23181(DIP2A) 55902(ACSS2) 84532(ACSS1)

glutamate - akg:
GLUDxm Glutamate Dehydrogenase (NAD), Mitochondrial glu_L[m] + h2o[m] + nad[m] <=> akg[m] + h[m] + nadh[m] + nh4[m]
  genes: 	2746(GLUD1) 2747(GLUD2)
AKGDm	2-oxoglutarate dehydrogenase Astrocyte  akg[mA] + coa[mA] + nad[mA] --> co2[mA] + nadh[mA] + succoa[mA]
  genes: (4967.2 and 1738.1 and 8050.1 and 1743.1) or (4967.1 and 1738.1 and 8050.1 and 1743.1)

gluthatione:
GLYOXm Hydroxyacylglutathione Hydrolase, Mitochondrial h2o[m] + lgt_S[m] -> gthrd[m] + h[m] + lac_D[m]
  genes: 	3029(HAGH)
GTHO Glutathione Oxidoreductase gthox[c] + h[c] + nadph[c] -> 2.0 gthrd[c] + nadp[c]
  genes: 2936(GSR)


```{r}
interesting_genes <- "BDHm|OCOAT1m|AACOAT|HMGLm|ACACT1rm|HMGCOASim|CSm|ACSm|GLUDxm|AKGDm|GLYOXm|GTHO"
H  %>% filter(str_detect(ID, regex(interesting_genes,ignore_case = T))) 
C %>% filter(str_detect(ID, regex(interesting_genes,ignore_case = T))) 
O %>% filter(str_detect(ID, regex(interesting_genes,ignore_case = T))) 
```



```{r}

H %>% writexl::write_xlsx("hyper_genes_rxns_df_aa.xlsx")
C %>%writexl::write_xlsx("central_genes_rxns_df_aa.xlsx")
O %>%writexl::write_xlsx("optimal_genes_rxns_df_aa.xlsx")

```



```{r}


H %>% dplyr::select(matches('gene')) %>% c %>% unlist %>% paste0%>% str_extract("\\d+") %>% na.omit() %>% as.character() -> H_only_genes
C%>% dplyr::select(matches('gene')) %>% c %>% unlist %>% paste0%>% str_extract("\\d+") %>% na.omit() %>% as.character() -> C_only_genes
O%>% dplyr::select(matches('gene')) %>% c %>% unlist %>% paste0%>% str_extract("\\d+") %>% na.omit() %>% as.character() -> O_only_genes


intersect(H_only_genes,C_only_genes )
intersect(H_only_genes,O_only_genes )
intersect(C_only_genes,O_only_genes )
```
































