---
title: "R Notebook"
output: html_notebook
---


```{r fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
library(clusterProfiler)
library(DOSE)
library(ReactomePA)
library(tidyverse)
library(readxl)

c('PGM', 'ACYP', 'PGI', 'PGK','PYK', 'HEX1', 'DPGase', 'TPI', 'PFK', 'ENO', 'GAPD', 'DPGM', 'FBA', 'G3PD2m' ,
'ACYP_Neuron', 'DPGM_Neuron', 'DPGase_Neuron', 'ENO_Neuron', 'FBA_Neuron', 'G3PD2m_Neuron', 'GAPD_Neuron', 'HEX1_Neuron', 'PFK_Neuron', 'PGI_Neuron', 'PGK_Neuron', 'PGM_Neuron', 'PYK_Neuron', 'TPI_Neuron' ,
'ATPS4m_Neuron', 'CYOOm2_Neuron', 'CYOR-u10m_Neuron', 'NADH2-u10m_Neuron', 'PPA_Neuron', 'PPAm_Neuron' ,
'PPAm', 'ATPS4m', 'CYOOm2', 'CYOR-u10m', 'NADH2-u10m', 'PPA', 'GLCt1r', 'GLCt1r_Neuron', 'GLCt2r') -> subsystems


filter_out_subsystems <- function(df){df[!(df$ID %in% subsystems),]}


 optimalhubs.xlsx <- read_excel('optimalhubs.xlsx') %>% filter(total > .5) %>% filter_out_subsystems
 centralhubs.xlsx <- read_excel('centralhubs.xlsx') %>% filter(Centrality > 1.10)%>% filter_out_subsystems
 hyperhubs.xlsx   <- read_excel('hyperhubs.xlsx')%>% filter(total > .5 | Centrality > 1.10)%>% filter_out_subsystems
```


```{r fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
get_genes_from_module <- function(hubs){     hubs %>% 
                                             dplyr::select(matches('gene')) %>% purrr::reduce(c) %>% na.exclude() %>% str_c(collapse = ", ") %>% 
                                            str_split(pattern=",") %>%str_extract_all('\\d+') %>% unlist() %>% str_trim %>% unique()}

hubs <- list(optimalhubs.xlsx, centralhubs.xlsx, hyperhubs.xlsx) %>% map(get_genes_from_module) %>% purrr::set_names(c("Optimal", "Central", "Hyper"))

 #setdiff(c(1,2,3,4),c(2,3,8))

c(setdiff(hubs$Optimal, hubs$Central))-> optimal_genes
c(setdiff(hubs$Central, hubs$Optimal))-> central_genes
unique(c(intersect(hubs$Optimal, hubs$Central),hubs$Hyper)) -> hyper_genes
intersect(central_genes, optimal_genes)
intersect(hyper_genes, optimal_genes)
intersect(hyper_genes, central_genes)    -> get_rid_from_hyper
setdiff(hyper_genes, get_rid_from_hyper) -> hyper_genes

rbind(
 optimalhubs.xlsx %>% dplyr::select(c('ID','Name','Reaction',"Node"),matches('gene')),
 centralhubs.xlsx  %>% dplyr::select(c('ID','Name','Reaction',"Node"),matches('gene')),
 hyperhubs.xlsx   %>% dplyr::select(c('ID','Name','Reaction',"Node"),matches('gene'))) -> all_hubs_info



```

ketone bodies:
BDHm    (R)-3-Hydroxybutanoate:NAD+ Oxidoreductase bhb[m] + nad[m] <=> acac[m] + h[m] + nadh[m]
  genes: 56898(BDH2) 622(BDH1)

OCOAT1m 3-Oxoacid Coa-Transferase acac[m] + succoa[m] -> aacoa[m] + succ[m]
 genes: 5019(OXCT1) 64064(OXCT2)


AACOAT  Acetoacetyl Coenzyme A:Acetate Coa-Transferase acac[c] + atp[c] + coa[c] -> aacoa[c] + amp[c] + ppi[c]:
 genes: 65985(AACS)


HMGLm   Hydroxymethylglutaryl Coenzyme A Lyase hmgcoa[m] -> acac[m] + accoa[m]
  genes: 	3155(HMGCL) 54511(HMGCLL1)

ACACT1rm Acetyl Coenzyme A C-Acetyltransferase, Mitochondrial 2.0 accoa[m] -> aacoa[m] + coa[m]

HMGCOASim Hydroxymethylglutaryl Coenzyme A Synthase (Ir) aacoa[m] + accoa[m] + h2o[m] -> coa[m] + h[m] + hmgcoa[m]
  genes: 3157(HMGCS1) 3158(HMGCS2)

CSm Citrate Synthase accoa[m] + h2o[m] + oaa[m] -> cit[m] + coa[m] + h[m]
 genes: 	1431(CS)

ACSm Acetyl Coenzyme A Synthetase ac[m] + atp[m] + coa[m] -> accoa[m] + amp[m] + ppi[m]
	genes: 23181(DIP2A) 55902(ACSS2) 84532(ACSS1)

glutamate - akg:
GLUDxm Glutamate Dehydrogenase (NAD), Mitochondrial glu_L[m] + h2o[m] + nad[m] <=> akg[m] + h[m] + nadh[m] + nh4[m]
  genes: 	2746(GLUD1) 2747(GLUD2)
AKGDm	2-oxoglutarate dehydrogenase Astrocyte  akg[mA] + coa[mA] + nad[mA] --> co2[mA] + nadh[mA] + succoa[mA]
  genes: (4967.2 and 1738.1 and 8050.1 and 1743.1) or (4967.1 and 1738.1 and 8050.1 and 1743.1)

gluthatione:
GLYOXm Hydroxyacylglutathione Hydrolase, Mitochondrial h2o[m] + lgt_S[m] -> gthrd[m] + h[m] + lac_D[m]
  genes: 	3029(HAGH)
GTHO Glutathione Oxidoreductase gthox[c] + h[c] + nadph[c] -> 2.0 gthrd[c] + nadp[c]
  genes: 2936(GSR)


```{r}
interesting_genes <- "BDHm|OCOAT1m|AACOAT|HMGLm|ACACT1rm|HMGCOASim|CSm|ACSm|GLUDxm|AKGDm|GLYOXm|GTHO"
all_hubs_info  %>% filter(str_detect(ID, regex(interesting_genes,ignore_case = T)))
```

```{r fig.height=4, fig.width=14, message=FALSE, warning=FALSE}


hubs <- list(optimal_genes, central_genes, hyper_genes) %>% purrr::set_names(c("Optimal", "Central", "Hyper"))

compareCluster(hubs, fun='enrichPathway', pvalueCutoff=1e-3) -> clusters_plot
clusters_plot %>% dotplot
```
```{r fig.height=4, fig.width=14, message=FALSE, warning=FALSE}

compareCluster(hubs, fun='enrichKEGG', pvalueCutoff=1e-2) -> clusters_plot
clusters_plot %>% dotplot
```

```{r fig.height=4, fig.width=14, message=FALSE, warning=FALSE}

compareCluster(hubs, fun='enrichGO', pvalueCutoff=0.001, OrgDb= org.Hs.eg.db )  -> clusters_plot
clusters_plot %>% dotplot
```

```{r}

get_rxns_from_genes <- function(a_gene){all_hubs_info %>% filter(str_detect(VMH_gene_IDs,a_gene) | str_detect(KEGG_genes,a_gene))%>%
                                         dplyr::select(c('ID','Name','Reaction',"Node"))  %>% mutate(Gene = a_gene)}

map_df(hubs$Hyper,get_rxns_from_genes)  %>% write_csv("hyper_genes_rxns_df.csv")
map_df(hubs$Central,get_rxns_from_genes) %>%write_csv("central_genes_rxns_df.csv")
map_df(hubs$Optimal,get_rxns_from_genes) %>%write_csv("optimal_genes_rxns_df.csv")


```

```{r}
intersect(hubs$Hyper,hubs$Central)
intersect(hubs$Hyper,hubs$Optimal)
intersect(hubs$Optimal,hubs$Central)
```


```{r}
interest <- clusters_plot@compareClusterResult %>% filter(Cluster=="Hyper")  %>% .[["geneID"]] %>% str_split(pattern="/") %>% unlist()

map_df(interest,get_rxns_from_genes)

```


